simulation & inference of transients

### 2019-07-17 notebook/may-outbreak.Rmd
exploring

```{r message=FALSE}
library(MDPtoolbox)
library(sarsop) # remotes::install_github("boettiger-lab/sarsop")
library(tidyverse) # for plotting
library(mdplearning)
```

@@

states <- seq(0,2, length=100)
actions <- states
observations <- states
sigma_g <- 0.05
sigma_m <- 0.0

r <- 0.8
K <- 1.53
q <- 2
b <- .2
eps <- states[2]
Tmax <- 100

may <- function(a){  
  function(x, h){ # May
    x <- pmax(x - h, 0)   # harvest then recruit   
    x + x * r * (1 - x / K)  - a * x ^ q / (x ^ q + b ^ q) + eps
  }
}

@@

```{r}
a <- 0.29
m <- fisheries_matrices(states, actions, observations, reward_fn, 
                        may(a), sigma_g, sigma_m, noise = "lognormal")
```

```{r, results="hide"}
soln <- mdp_value_iteration(m$transition, m$reward, discount)
```

```{r}
df <- tibble(state = states,
             action = actions[soln$policy],
             value = soln$V)
## tipping point at maximum action:
df %>% ggplot(aes(state, action)) + geom_point() 
```

```{r}
x0 <- which.min(abs(states - 0.1))
#x0 <- which.min(abs(states - 2))
df <- mdp_planning(m$transition, m$reward, discount, model_prior = c(1), 
                   policy = soln$policy, x0 = x0, Tmax = 100)
df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

@@

```{r}
possible_a <- seq(.25, .34, by = 0.005)

true_a <- 0.29
true_i <- which.min(abs(possible_a - true_a))

@@

@@

## Small a assumption

```{r}
mu <- possible_a[true_i - 8]
prior <- dnorm(possible_a, mu, 0.01)
prior <- prior / sum(prior)
```

```{r}
barplot(prior)
f <- may(mu)
tibble(x = states[1:80],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + geom_line() +
  geom_point() +  geom_hline(aes(yintercept = 0))
```

@@

## large a assumption

```{r}
mu <- possible_a[true_i + 8]
prior <- dnorm(possible_a, mu, 0.01)
prior <- prior / sum(prior)
```

```{r}
barplot(prior)
f <- may(mu)
tibble(x = states,
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + geom_line() +
  geom_point() +  geom_hline(aes(yintercept = 0))
```

## Perfect knowledge

```{r}
mu <- possible_a[true_i]
#prior <- dnorm(possible_a, mu, 0.01)
#prior <- prior / sum(prior)
prior <- numeric(length(possible_a))
prior[true_i] <- 1
```

```{r}
barplot(prior)
f <- may(mu)
tibble(x = states,
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + geom_line() +
  geom_point() +  geom_hline(aes(yintercept = 0))
```

```{r}
sim$df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

>>>>

```{r message=FALSE}
library(MDPtoolbox)
library(sarsop) # remotes::install_github("boettiger-lab/sarsop")
library(tidyverse) # for plotting
library(mdplearning)

#if(interactive()) ggplot2::theme_set(ggdark::dark_theme_light())
#ggplot2::theme_set(theme_grey())
#ggdark::invert_geom_defaults()
```

@@

states <- seq(0,2, length=200)
actions <- states
observations <- states
sigma_g <- 0.05
sigma_m <- 0.0

r <- 0.8
K <- 1.53
q <- 2
b <- .2
eps <- states[2]/10
Tmax <- 100

may <- function(a){  
  function(x, h){ # May
    # x <- pmax(x - h, 0)   # harvest then recruit   
    x + x * r * (1 - x / K)  - a * x ^ q / (x ^ q + b ^ q) + eps - h
  }
}

@@

@@

```{r}
possible_a <- seq(.25, .34, by = 0.005)
true_a <- 0.26   ## reality has just a transient.  use  0.277 for stronger ghost
believe_a <- 0.31 ## believe there's an attractor
true_i <- which.min(abs(possible_a - true_a))
```

```{r}
f <- may(true_a)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + geom_line() +
  geom_point() +  geom_hline(aes(yintercept = 0))
```

```{r}
f <- may(believe_a)
tibble(x = states[1:120],
       f = f(x,0) - x) %>%
  ggplot(aes(x, f)) + geom_line() +
  geom_point() +  geom_hline(aes(yintercept = 0))
```

```{r}
m_true <- fisheries_matrices(states, actions, observations, reward_fn, 
                        may(true_a), sigma_g, sigma_m, noise = "lognormal")

m <- fisheries_matrices(states, actions, observations, reward_fn, 
                        may(believe_a), sigma_g, sigma_m, noise = "lognormal")
```

```{r, results="hide"}
soln <- mdp_value_iteration(m$transition, m$reward, discount)

opt_soln <- mdp_value_iteration(m_true$transition, m_true$reward, discount)

@@

Policy based on the belief (i.e. that system is bi-stable)

```{r}
df <- tibble(state = states,
             action = actions[soln$policy],
             value = soln$V)
df %>% ggplot(aes(state, action)) + geom_point() 
```

Optimal policy (knowing it is really just a transient)
```{r}
tibble(state = states,
       action = actions[opt_soln$policy]) %>% 
  ggplot(aes(state, action)) + geom_point() 
```

```{r}
no_policy <- numeric(length(states)) + 1

df <- mdp_planning(m_true$transition, m_true$reward, discount, model_prior = c(1), 
                   policy = no_policy, x0 = x0, Tmax = 100)

df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

Result experienced by incorrect belief:  initial in-action followed by need for continued maintenance to prevent high-level outbreak:

df <- mdp_planning(m_true$transition, m_true$reward, discount, model_prior = c(1), 
                   policy = soln$policy, x0 = x0, Tmax = 100)

df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

Expected result based on the belief: stable low level is acceptable, so no action is required:

df <- mdp_planning(m$transition, m$reward, discount, model_prior = c(1), 
                   policy = soln$policy, x0 = x0, Tmax = 100)

df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

Optimal strategy knowing this is just a transient:

df <- mdp_planning(m_true$transition, m_true$reward, discount, model_prior = c(1), 
                   policy = opt_soln$policy, x0 = x0, Tmax = 400)

df %>% mutate(state = states[state], action = actions[action]) %>% 
  ggplot(aes(time, state)) + geom_point()+geom_path() + 
  geom_line(aes(time, action), col="blue")
```

@@

## a near ghost

```{r}
prior <- dnorm(possible_a, believe_a, 0.03)
prior <- prior / sum(prior)
```

```{r}
data.frame(a = possible_a, probability = prior) %>%
  ggplot(aes(a,prior)) + geom_bar(stat="identity") +
  geom_vline(aes(xintercept = true_a), col="red", lwd=1, lty=2) + 
  geom_vline(aes(xintercept = believe_a), col="blue", lwd=1, lty=2) 
```

@@
